\documentclass[11pt]{article}
\usepackage{quanttheme}

\newcolumntype{Y}{>{\raggedright\arraybackslash}X}
\newcommand{\tk}[1]{\texttt{#1}}

\begin{document}
\thispagestyle{empty}

\titlebanner
  {QuantAnomalies}
  {Quantitative Anomalies as Market Opportunities}
  {Progress Presentation \quad\textbullet\quad Statistical \& AI Dashboard for Equity and Forex Markets}

\vspace{1.2em}

\section{Title of the Study}
\textbf{Quantitative Anomalies as Market Opportunities — A Statistical \& AI
Dashboard for Equity and Forex Markets.}

\begin{opp}
In plain terms: textbook finance says markets are ``efficient'' — prices already
reflect everything, risk never changes, and nothing is predictable. Reality is
messier, and those cracks in the theory are recurring patterns called
\textbf{anomalies}. Our project hunts for those anomalies and reframes them as
\textbf{opportunities}: places where the odds may tilt in a careful observer's favour.
\end{opp}

\section{Objectives of the Study}
Each objective targets one well-known market anomaly and turns it into a usable
opportunity signal.

\begin{enumerate}
\item \textbf{Macro-driver opportunity.} Measure \emph{which} macroeconomic forces
(interest rates, inflation, the fear index, oil, gold, the dollar) move an asset's
returns — and crucially with \emph{what time delay}. If a driver acts weeks later,
that lag is a window to anticipate moves.

\item \textbf{Risk-regime opportunity.} Model how risk itself rises and falls over
time (volatility clustering) and detect that crashes happen far more often than a
bell curve predicts (fat tails). Spotting a ``stormy'' regime early is a chance to
cut risk — or to recognise unusually cheap calm.

\item \textbf{Mean-reversion opportunity.} Find pairs of currencies whose prices are
tethered over the long run (cointegration), then flag when they drift unusually far
apart — history says the gap tends to snap back. This is the basis of
market-neutral ``statistical arbitrage''.

\item \textbf{Opportunity synthesis (the recommendation engine).} Combine all of the
above into one automated scanner that ranks what is unusual or actionable
\emph{right now}, and uses a \textbf{local AI analyst} to explain each finding in
plain English.

\item \textbf{Accessibility \& robustness.} Make it work for \emph{any} ticker the
user types, on a reliable data pipeline, presented through a clear, interactive
dashboard.
\end{enumerate}

\section{Motivation for Selecting the Topic}
\begin{itemize}
\item \textbf{The anomalies are real and well-studied, but rarely made tangible.}
Lagged macro effects, volatility clustering, fat tails, and cointegration all appear
in finance literature; we wanted to make them \emph{visible and actionable} rather
than abstract equations.
\item \textbf{Reframing risk as opportunity.} The same statistics used to warn about
risk can also point to opportunity — we wanted a tool that does both.
\item \textbf{Bridging statistics and modern AI.} We combine classic econometrics
(regression, GARCH, cointegration) with a \textbf{local large language model} that
turns numbers into a readable analyst note — private, free to run, and on our own GPU.
\item \textbf{Learning value.} The project spans real statistical modelling, live
financial-data engineering, full-stack web development, and on-device AI.
\end{itemize}

\section{Data \& APIs Used}
\textit{In place of a questionnaire and a survey sample: our ``instrument'' is a set
of financial-data APIs, and our ``sample'' is the live market history they return.}

\subsection{A. External data-source APIs (collecting raw market data)}
\begin{itemize}
\item \textbf{Yahoo Finance} (via the \tk{yfinance} library) — daily price history
(2015--2025) for any equity, index, currency pair, future, or crypto. Examples:
\tk{\textasciicircum{}GSPC} (S\&P 500), \tk{\textasciicircum{}VIX} (volatility index),
\tk{CL=F} (oil), \tk{GC=F} (gold), \tk{\textasciicircum{}TNX} (10-year Treasury
yield), \tk{DX-Y.NYB} (US dollar index), and 45 forex pairs.
\item \textbf{FRED — U.S. Federal Reserve Economic Data} — macroeconomic series:
CPI (inflation), Fed Funds Rate, and Unemployment. A free mirror, \textbf{DBnomics},
is a built-in backup so a single outage never stops the study.
\end{itemize}

\subsection{B. Our own analysis API (FastAPI backend)}
Turns raw data into results:
\begin{itemize}
\item \tk{/api/macro-regression/*} — regression, Granger-causality, correlation, time series
\item \tk{/api/garch/*} — volatility model, clustering tests, return distribution
\item \tk{/api/pairs/*} — cointegration tests, best-pair spread \& signals, correlation
\item \tk{/api/recommendations} — the ranked opportunity scan
\item \tk{/api/llm/info} — status of the local AI analyst
\end{itemize}

\subsection{C. Local AI API}
An on-device \tk{llama.cpp} server (OpenAI-compatible) running the \textbf{Qwen3-4B}
model on our \textbf{NVIDIA RTX 4050 GPU}, used to explain detected opportunities in
plain language. No data leaves the machine.

\section{Pilot Study Conducted to Date}
We have a \textbf{working end-to-end prototype}, validated on real data.

\begin{keybox}
\textbf{Data pipeline (validated).} Live downloads work for all sources; a corrupted
early cache (the volatility and oil series had been mixed up) was found and fixed.
Every series now sits in its correct real-world range — e.g. VIX 9--83, oil $-37$ to
124, 10-year yield 0.5--4.9\%.
\end{keybox}

\textbf{Three analysis pillars — working, with sample findings on the S\&P 500:}
\begin{itemize}
\item \emph{Macro:} macro factors explain about \textbf{64\%} of monthly return
variation (R² ≈ 0.64); the model surfaces which factors are significant and at what lag.
\item \emph{Volatility:} GARCH ``persistence'' ≈ \textbf{0.9955} (shocks fade very
slowly) and excess kurtosis ≈ \textbf{16} (strong fat tails) — risk is \emph{not}
constant and crashes are under-estimated by the normal curve.
\item \emph{Pairs:} among major forex pairs, \textbf{USDCHF/USDJPY} tested
cointegrated (p ≈ 0.008) with a mean-reversion half-life around 64 days.
\end{itemize}

\textbf{Recommendation engine (v1 working).} Four detectors (volatility regime, tail
move, trend, forex mean-reversion) rank opportunities by severity and assign a
confidence score. The local AI produces a grounded plain-English note in
≈4 seconds on the GPU (≈28 tokens/sec), citing only the real detected numbers.

\textbf{Dashboard (working).} Interactive web app: search any ticker, pick forex
pairs, switch time ranges, and read explained results across five tabs (Overview,
Opportunities, Macro, GARCH, Pair Trading).

\begin{opp}
\textbf{In short:} the pilot confirms the data, the statistics, the AI-explanation
layer, and the user interface all work together on live markets. Next steps
(standardised comparisons, an options / Black–Scholes module, and a back-test of the
signals) are tracked in the project roadmap.
\end{opp}

\section{Modules Built So Far}

\subsection*{Backend (Python / FastAPI)}
\begin{itemize}
\item \tk{data\_loader.py} — data acquisition (Yahoo Finance, FRED, DBnomics), caching, dataset builders
\item \tk{analysis/macro\_regression.py} — Pillar 1: OLS lag regression, Granger causality, correlation
\item \tk{analysis/garch.py} — Pillar 2: GARCH(1,1), volatility clustering, return distribution
\item \tk{analysis/pairs.py} — Pillar 3: cointegration, spread / z-score, trading signals
\item \tk{analysis/recommender.py} — the opportunity engine: detectors, ranking, confidence
\item \tk{llm\_client.py} — provider-agnostic local-LLM client (GPU server + CPU fallback)
\item \tk{main.py} — FastAPI endpoints tying it all together
\end{itemize}

\subsection*{Frontend (React + Vite)}
\begin{itemize}
\item Five tab panels — Overview, Opportunities, Macro Regression, GARCH Volatility, Pair Trading
\item Shared components — SVG icon set, info-tooltips, time-range filter, ticker search, forex selector, status states, animated background
\item Support — chart theme, time-range utilities, data-fetch hook, API client
\end{itemize}

\section{Future Plans — Expanding the Opportunity Engine}
Each future module is another \emph{lens} for spotting opportunity. The engine's
principle stays the same: \textbf{statistics detect, the AI explains}.

\subsection*{1. Options \& Black–Scholes}
Price European call / put options and their Greeks, and back out the market's
\textbf{implied volatility}. \textbf{Opportunity:} compare implied volatility with
our GARCH forecast — when the market pays far more (or less) for risk than the model
expects, options are ``rich'' or ``cheap''.

\subsection*{2. Valuation — ``is the stock bloated / overpriced?''}
Pull fundamental ratios (P/E, P/B, P/S, PEG, EV/EBITDA) and compare them with the
stock's own history and its peers, alongside a technical ``stretch'' check (how far
price sits above its 200-day average, and RSI). \textbf{Opportunity:} flag names
``priced for perfection'' (avoid / potential short) versus those ``on sale''
(potential value buys).

\subsection*{3. More opportunity lenses}
\begin{itemize}
\item \textbf{Momentum \& relative strength} — rank a universe by risk-adjusted
12–1 month momentum to spot rotation leaders and laggards.
\item \textbf{Oversold bounce} — distance from the 52-week high plus RSI to find
mean-reversion candidates.
\item \textbf{Volatility squeeze} — detect compression that often precedes a breakout.
\item \textbf{Correlation-regime shift} — assets that usually move together
decoupling (a diversification or pairs opportunity).
\item \textbf{Seasonality / calendar effects} — turn-of-month and ``sell in May''
style documented anomalies.
\item \textbf{Alpha vs beta} — separate market-driven moves from stock-specific mispricing.
\item \textbf{Unusual volume} — accumulation / distribution versus the average.
\item \textbf{Risk budgeting} — turn the detected regime into a suggested position size.
\end{itemize}

\section{Decision Engine: Rules-based vs LLM-based}
We support \textbf{both}, and today combine them: \emph{rules detect, the LLM
explains}. A toggle lets the user choose the mode; pure rules is the safe default.

\begin{table}[h]
\renewcommand{\arraystretch}{1.3}
\begin{tabularx}{\linewidth}{>{\bfseries}l Y Y}
\toprule
& \textbf{\color{qnavy}Rules-based} & \textbf{\color{qnavy}LLM-based} \\
\midrule
Determinism & Fully deterministic, reproducible & Varies with settings \\
Explainability & Transparent thresholds & Natural-language reasoning \\
Numbers & Exact by construction & Must be guarded against errors \\
Flexibility & Rigid; new rules need code & Synthesises novel combinations \\
Speed \& cost & Instant, free & ≈4 s on GPU, compute cost \\
Works offline & Yes & Needs the local model \\
Best at & Detection \& signals & Explanation \& synthesis \\
\bottomrule
\end{tabularx}
\end{table}

\begin{opp}
Our design uses each for what it is good at — rules produce the trustworthy numbers,
the LLM produces the readable story. A future \textbf{LLM decision-maker} mode would
let the model also weigh and rank signals (not just explain), kept honest by
rule-based guardrails and number validation.
\end{opp}

\section*{Appendix — Each anomaly in one line (layman)}
\begin{itemize}
\item \textbf{Lagged macro effect:} \textit{News ripples through markets over weeks,
not instantly — know the lever and the delay, and you can position ahead.}
\item \textbf{Volatility clustering:} \textit{Calm days cluster, and so do wild days —
storms arrive in groups, giving you warning to manage risk.}
\item \textbf{Fat tails:} \textit{Extreme crashes happen much more often than the
textbook bell curve says — plan for them.}
\item \textbf{Cointegration / pairs:} \textit{Two currencies on one leash: when they
drift far apart, they usually snap back — bet on the gap closing, up market or down.}
\end{itemize}

\end{document}
